---
title: "native_food_preferences"
author: "Kevin Havis"
format: html
---

```{r}
source('load_data.R')
```

1.  Do better educated U.S. adults have greater knowledge of food cuisines?

Independent variable = `education`

Dependent variable = `knowledge_cuisines`

Analysis method: ANOVA



```{r}
# Calculate descriptive statistics
prop.table(table(data$education, data$knowledge_cuisines))

data %>%
  drop_na(education, knowledge_cuisines) |> 
  count(education, knowledge_cuisines) %>%
  group_by(education, knowledge_cuisines) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup()

data %>%
  select(education, knowledge_cuisines) |> 
  drop_na(education, knowledge_cuisines) %>%
  count(education, knowledge_cuisines) %>%
  group_by(education) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup()
```



2.  Do different generations (age groups) have different preferences of cuisines?

ANOVA across cuisines 



3.  What cuisines do women prefer compared to men?

Difference in means for each country, divided by gender




